Multiresolution Neural Field Estimation

### 2011-02-28 Multiresolution_framework/ltx/manuscript.tex
edits to paper

\author{In no particular order: \parham{P. Aram}, \mike{M. Dewar}, \dean{D.R. Freestone}, \ken{K.Scerri}, D.B. Grayden and V. Kadirkamanathan % <-this % stops a space
\thanks{P. Aram* and V. Kadirkamanathan are with the Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, S1 3JD, U.K. (e-mail:p.aram@sheffield.ac.uk; visakan@sheffield.ac.uk).}% <-this % stops a space
\thanks{M. Dewar is with the School of Informatics, University of Edinburgh, Edinburgh, EH8 9AB, U.K. (e-mail:mike.dewar@inf.ed.ac.uk)}}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

@@

\begin{abstract}
%\boldmath
% The Integro Difference Equation (IDE) is an increasingly popular model of spatio-temporal processes. Here we develop a Multiresolution Approximation (MRA) framework for the IDE based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is approached in a Maximum Likelihood (ML) framework using the Expectation Maximisation (EM) algorithm. Examples are given to demonstrate the features of the model.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Neural field model, multiresolution approximation (MRA), expectation maximization (EM) algorithm, wavelets.
\end{IEEEkeywords}

@@

We have a clear definition and where to stimulate and how to stimulate.\\
\cut{This color is used to indicate bits we can get rid of if we dont have enough space, use the command: \textbackslash cut$\left\lbrace  \right\rbrace$}

\section{Introduction}
\IEEEPARstart{T}{his} paper introduces a multiresolution data-driven framework for neural field modeling. Neural field models describe the meso and macroscopic dynamics of the human brain. Modeling the brain at this level is extremely important, since diseases such as epilepsy and \parham{Parkinson's disease} are manifested at this scale through pathological oscillations. The ability to create patient-specific neural field models will not only contribute to our knowledge base of poorly understood diseases such as epilepsy, but may also enable the development of new treatment strategies. This is particularly poignant with the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices using corrective or modulating electrical stimulation operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using systems theory. The design of a stimulation strategy using systems theory requires a model. Neural mass and neural field models are the ideal candidates for estimation and control algorithms. It is expected that parameters of the neural fields models will be patient specific and will therefore need to be inferred from data.

@@

This paper is an extension to the framework presented Freestone et al. (2011), where a finite element model of the neural field (via a global Galerkin projection) was formed, using a basis function decomposition, to transform the PDE neural field equations into a finite dimension system to facilitate state and parameter estimation. The basis function decomposition in this previous study did not account for all of the spatial dynamics. Thus, it can be improved.. In this current paper, a multiresolution approach is taken...

@@

\section{Integro-Difference Equation Neural Field Model}
The stochastic integro-difference equation (IDE) form of the Amari neural field  formulation~\cite{Amari1977} is given by (see \dean{cite NeuroImage paper} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f\left(.\right)$. \dean{The activation function is of the form $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r}) + v_0$}. The term $e_t(\mathbf r)$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$, where $\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$  denotes a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. For simplicity, \cut{we denote the current time step by $t$ and the future time step by $t+1$.} The observation equation describing the electrophysiological recordings is given by 
\begin{equation}\label{eq:ObservationEquation}
	\mathbf y_t = \int_{\Omega} { m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t, 
\end{equation}
where $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$, is corrupted by an i.i.d normally distributed zero-mean white noise $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$ with $\mathbf{\Sigma}_{\epsilon}=\sigma_{\epsilon}^2\mathbf I_{n_y} $. The superscript $\top$ denotes the transpose operator.

@@

% \section{B-spline wavelet IDE model}
The multiresolution approximation can be implemented using various classes of scaling and wavelet functions. B-spline functions were chosen due their compact support and the ability to analytically calculate convolution and inner product to form the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how the matrices are formed, the $m^{th}$ order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 
\begin{align}
N_{m}\left(r\right)=&\left(N_{m-1}\ast N_{1}\right)\left(r\right)\nonumber \\
=&\int_0^{1} N_{m-1}\left( r-r'\right)dr' \quad m>1,
\label{SplineConvolutionIntegral}
\end{align}
where $\ast$ denotes convolution and $N_1\left(s\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Eq.~\ref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ times convolution of indicator function with itself since
\begin{equation}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is (see \cite{Chui1992} for proof) 
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. The $m^{th}$ order B-spline wavelet function is defined as (see \cite{Chui1992}) 
\begin{align}
 \varphi & _{m}\left(r\right) = \sum_{n=0}^{3m-2} q_n N_{m}\left(2r-n\right) \\
 q & _n = \frac{\left(-1\right)^n}{2^{m-1}} \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$},
\end{align}
where $q_n$ are the coefficients. By exploiting properties Eqs.~\ref{eq:BsplineConvolution} and \ref{eq:BsplineInnerProduct} the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta} can be computed. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$ the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale.  In this paper $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta}.

\section{State and Parameter Estimation}

The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used to in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Gibsona2005,Roweis1999,Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. For a multivariate linear dynamic system, the bound used is a quadratic of the form
\begin{equation}
 \mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
which has a maximum at
\begin{align}
 \boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0^\top-\boldsymbol\upsilon_1^\top),
\end{align}
where
\begin{align}\label{eq:upsilon0}
 \boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\\ 
 \boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\label{eq:upsilon1} \\
 \boldsymbol\Upsilon&=T_s\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\mathbf{U}^{\mathsf T} \boldsymbol\Lambda_{x}^{-1}\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf{U}]_{j,i}\label{eq:Upsilon}.
\end{align}
The $\left(i,j\right)^{th}$ block of the block matrix $\mathbf U_{n_x \times n_x}$ is given by
\mike{you need to define and sort out the $[]_{ij}$ notation!}
\begin{align}
\left[ \mathbf U\right] _{i,j}&=\int_{\boldsymbol \Omega}\left[\mu(r) \right]_i \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(\mathbf{r-r'}\right) dr'\right]_{j:} dr.
\end{align}
where $[.]_{j:} $ denotes $j^{th}$ row of the matrix.\cut{The second terms in Eqs. \ref{eq:upsilon0}, \ref{eq:upsilon1}, and \ref{eq:Upsilon} can be computed as a once-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the method in \cite{Dewar2009}.} The matrices $\Xi_0$ and $\Xi_1$ are calculated using the Rauch Tung Streibel (RTS)\cite{RAUCH1965} smoother outputs: state estimates, $\hat{\mathbf x}_t$, covariance, $\mathbf P_t=\mathrm{cov}(\mathbf{x_t})$, and cross-covariance matrix $\mathbf M_t=\mathrm{cov}(\mathbf{x}_t,\mathbf{x}_{t-1})$ 
\begin{align}\label{eq:Xivariables}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).
\end{align}

@@

\section{Results}
To demonstrate the performance of the MRA estimation framework, data was generated synthetically  using Eqs.~\ref{eq:DiscreteTimeModel} and \ref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 100 realizations of 1 second of data were generated where $\boldsymbol\Sigma_v=0.1 \times \mathbf{I}_{n_y}$ and $\eta(r-r')=\phi_{3,-2}(r-r')$. The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms respectively. The distance between adjacent sensors was $0.5$~mm resulting into 161 observations. \parham{The observation kernel is modeled by $\phi_{4,k}(r-r')$} \dean{?}, providing a width of 0.8 mm at half the maximum amplitude of the observation kernel, allowing for some cross talk in the recordings. The actual connectivity kernel and its decomposition are plotted in \figurename{\ref{fig:KernelAndFreqResponse}} (a).

>>>>

\author{In no particular order: \parham{P. Aram}, \mike{M. Dewar}, \dean{D.R. Freestone}, \ken{K.Scerri}, D.B. Grayden and V. Kadirkamanathan % <-this % stops a space
\thanks{P. Aram* and V. Kadirkamanathan are with the Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, S1 3JD, U.K. (e-mail:p.aram@sheffield.ac.uk; visakan@sheffield.ac.uk).}% <-this % stops a space
\thanks{M. Dewar is with the School of Informatics, University of Edinburgh, Edinburgh, EH8 9AB, U.K. (e-mail:mike.dewar@inf.ed.ac.uk)}
\thanks{D.\ R.\ Freestone and D.\ B.\ Grayden are with the Department
of Electrical and Electronic Engineering, The University of Melbourne, VIC 3010, Australia {\tt\small dfreestone@bionicear.org, grayden@unimelb.edu.au}}}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

@@

\begin{abstract}
%\boldmath
The Integro-difference equation (IDE) is an increasingly popular model of spatio-temporal processes. Here we develop a multiresolution approximation (MRA) framework for the IDE neural field equations based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is approached in a Maximum Likelihood (ML) framework using the Expectation Maximisation (EM) algorithm. Examples are given to demonstrate the framework.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Neural field model, multiresolution approximation (MRA), Expectation Maximization (EM) algorithm, wavelets.
\end{IEEEkeywords}

@@

We have a clear definition and where to stimulate and how to stimulate.\\
\cut{This color is used to indicate bits we can get rid of if we don't have enough space, use the command: \textbackslash cut$\left\lbrace  \right\rbrace$}

\section{Introduction}
\IEEEPARstart{T}{he} human brain has an inherent multiresolution architecture. Spatial scales of information processing and transfer range from individual proteins in synapses, to neurons, to networks of neurons. The multiresolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multiresolution data-driven framework for neural field modeling, the multiresolution approximate integrodifference equation (MRAIDE), to address this challenge. Neural field models typically describe the meso and macroscopic dynamics of the human brain, but are also descriptive of finer-scale neurodynamics.

The ability to create patient-specific neural models will contribute to our knowledge base of poorly understood diseases such as epilepsy, and furthermore, may also enable the development of new treatment strategies. This is particularly poignant with the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices using corrective or modulating electrical stimulation operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using systems theory. The design of a stimulation strategy using systems theory requires a model. Neural mass and neural field models are the ideal candidates for estimation and control algorithms. It is expected that parameters of the neural fields models will be patient specific and will therefore need to be inferred from data.

@@

This paper is an extension to the framework presented Freestone et al. (2011), where a finite element model of the neural field (via a global Galerkin projection) was formed, using a basis function decomposition, to transform the PDE neural field equations into a finite dimension system to facilitate state and parameter estimation. The basis function decomposition in this previous study did not account for multiresolution architecture and spatial dynamics of the human cortex. Thus, it can be improved.. In this current paper, a multiresolution approach is taken...

@@

\section{Integro-Difference Equation Neural Field Model}
The stochastic integro-difference equation (IDE) form of the Amari neural field  formulation~\cite{Amari1977} is given by (see \dean{cite NeuroImage paper} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The term $e_t(\mathbf r)$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$, where $\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$  denotes a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. For simplicity, \cut{we denote the current time step by $t$ and the future time step by $t+1$.} The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f\left(.\right)$. \dean{The activation function is of the form $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r}) + v_0$}.

The observation equation describing the electrophysiological recordings is given by 
\begin{equation}\label{eq:ObservationEquation}
	\mathbf y_t = \int_{\Omega} { m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t, 
\end{equation}
where $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$, is corrupted by an i.i.d normally distributed zero-mean white noise $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$ with $\mathbf{\Sigma}_{\epsilon}=\sigma_{\epsilon}^2\mathbf I_{n_y} $. The superscript $\top$ denotes the transpose operator.

@@

The multiresolution approximation can be implemented using various classes of scaling and wavelet functions. B-spline functions were chosen due their compact support and the ability to analytically calculate convolution and inner product to form the matrices $ \boldsymbol\Lambda_x$ and $\boldsymbol \Lambda_{\theta}$. To show how the matrices are formed, the $m^{th}$ order cardinal B-spline scaling function is defined by the recurrence relation \cite{Chui1992} 
\begin{align}
N_{m}\left(r\right)=&\left(N_{m-1}\ast N_{1}\right)\left(r\right)\nonumber \\
=&\int_0^{1} N_{m-1}\left( r-r'\right)dr' \quad m>1,
\label{SplineConvolutionIntegral}
\end{align}
where $\ast$ denotes convolution and $N_1\left(s\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Eq.~\ref{SplineConvolutionIntegral} can be rewritten as $(m-1)$ times convolution of indicator function with itself since
\begin{equation}
 N_{m}\left(r\right)=\underbrace{\left(N_{1}\ast N_{1}\ast \cdots \ast N_{1}\right)}_{m-1\quad \text{convolutions}}\left(r\right).
\end{equation}
Using the associativity property of convolution we have
\setlength{\arraycolsep}{0.0em}
\begin{align}\label{eq:BsplineConvolution}
N_{m}\left( r\right) \ast N_{m'}\left(r\right)&=\underbrace{\overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m-1 \quad \text{convolutions}}\left(r\right) \ast \overbrace{\left(N_{1} \ast \cdots \ast N_{1}\right)}^{m'-1\quad \text{convolutions}}}_{m+m'-1 \quad \text{convolutions}}\left(r\right)\nonumber\\
&=N_{m+m'}\left(r\right).
\end{align}
A direct consequence of Eq.~\eqref{eq:BsplineConvolution} is (see \cite{Chui1992} for proof) 
\begin{align}
 \left\langle N_{m}\left(r-l_{1}\right), N_{m'}\left(r-l_{2}\right)\right\rangle=&N_{m+m'}\left(m+l_{1}-l_{2}\right)\nonumber \\
=&N_{m+m'}\left(m'+l_{2}-l_{1}\right),
\label{eq:BsplineInnerProduct}
\end{align}
where $\left\langle \cdot,\cdot\right\rangle $ denotes the inner product. The $m^{th}$ order B-spline wavelet function is defined as (see \cite{Chui1992}) 
\begin{align}
 \varphi & _{m}\left(r\right) = \sum_{n=0}^{3m-2} q_n N_{m}\left(2r-n\right) \\
 q & _n = \frac{\left(-1\right)^n}{2^{m-1}} \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$},
\end{align}
where $q_n$ are the coefficients. By exploiting properties Eqs.~\ref{eq:BsplineConvolution} and \ref{eq:BsplineInnerProduct} the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta} can be computed. In order to calculate elements of $\boldsymbol\Lambda_{x}$ and $\boldsymbol\Lambda_{\theta}$ the scaling and wavelet basis functions must be expanded in terms of $N_m$ at the appropriate scale.  In this paper $4^{th}$ order cardinal B-spline scaling and wavelet functions are used. Therefore, the $8^{th}$ and $12^{th}$ order B-spline values at integer points are required (see \cite{Goswami1999}) to compute the integrals in Eqs.~\ref{eq:Lambdax} and \ref{eq:Lambdatheta}.

\section{State and Parameter Estimation}
The state-space representation of the MRAIDE allows the use of the well known Expectation Maximization (EM) algorithm \cite{Dempster1977} to infer both the kernel and the field from electrophysiological data. The EM algorithm, when used to in this context \cite{Dewar2009}, yields the maximum likelihood kernel estimate and the posterior distribution of the field over time. We use the standard EM algorithm for linear dynamic systems \cite{Gibsona2005,Roweis1999,Shumway2000} and hence only describe aspects of the algorithm that are specific to MRA neural field equations. EM essentially finds increasingly tighter lower bounds on the likelihood of the kernel $p(\mathbf{y}_1 \ldots \mathbf{y}_T|w)$ so that, at convergence, the maximum of the bound corresponds to the maximum of the likelihood. For a multivariate linear dynamic system, the bound used is a quadratic of the form
\begin{equation}
 \mathcal Q\left(\boldsymbol \theta,\boldsymbol\theta'\right)=\beta+2T_s\left(\boldsymbol\upsilon_0-\boldsymbol\upsilon_1\right)\boldsymbol\theta-T_s\boldsymbol\theta^\top\boldsymbol\Upsilon\boldsymbol\theta,
\end{equation}
which has a maximum at
\begin{align}
 \boldsymbol \theta= \boldsymbol\Upsilon^{-\top}(\boldsymbol\upsilon_0^\top-\boldsymbol\upsilon_1^\top),
\end{align}
where
\begin{align}\label{eq:upsilon0}
 \boldsymbol\upsilon_0&=\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_0]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\\ 
 \boldsymbol\upsilon_1&=\xi\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf U]_{j,i}\label{eq:upsilon1} \\
 \boldsymbol\Upsilon&=T_s\sum_{i,j=1}^{n_x}[\boldsymbol\Xi_1]_{i,j}[\mathbf{U}^{\mathsf T} \boldsymbol\Lambda_{x}^{-1}\Sigma_{w}^{-1}\boldsymbol\Lambda_{x}^{-1}\mathbf{U}]_{j,i}\label{eq:Upsilon}.
\end{align}
The $\left(i,j\right)^{th}$ block of the block matrix $\mathbf U_{n_x \times n_x}$ is given by
\mike{you need to define and sort out the $[]_{ij}$ notation!}
\begin{align}
\left[ \mathbf U\right] _{i,j}&=\int_{\boldsymbol \Omega}\left[\mu(r) \right]_i \left[\int_{\boldsymbol\Omega} \boldsymbol\mu\left(r'\right)\boldsymbol \lambda^\top \left(\mathbf{r-r'}\right) dr'\right]_{j:} dr.
\end{align}
where $[.]_{j:} $ denotes $j^{th}$ row of the matrix.\cut{The second terms in Eqs. \ref{eq:upsilon0}, \ref{eq:upsilon1}, and \ref{eq:Upsilon} can be computed as a once-off before the commencement of the EM iterations, which increases the speed of the M-step significantly compared to the method in \cite{Dewar2009}.} The matrices $\Xi_0$ and $\Xi_1$ are calculated using the Rauch Tung Streibel (RTS)\cite{RAUCH1965} smoother outputs: state estimates, $\hat{\mathbf x}_t$, covariance, $\mathbf P_t=\mathrm{cov}(\mathbf{x_t})$, and cross-covariance matrix $\mathbf M_t=\mathrm{cov}(\mathbf{x}_t,\mathbf{x}_{t-1})$ 
\begin{align}\label{eq:Xivariables}
\boldsymbol\Xi_0&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right) \\
 \boldsymbol\Xi_1&=\sum_{t=0}^{T-1}\left(\mathbf P_t+\mathbf{\hat x}_t\mathbf{\hat x}_t^\top\right).
\end{align}

@@

\section{Results}
To demonstrate the performance of the MRA estimation framework, data was generated synthetically  using Eqs.~\ref{eq:DiscreteTimeModel} and \ref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 100 realizations of 1 second of data were generated where $\boldsymbol\Sigma_v=0.1 \times \mathbf{I}_{n_y}$ and $\eta(r-r')=\phi_{3,-2}(r-r')$. The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms respectively. The distance between adjacent sensors was $0.5$~mm resulting into 161 observations. \parham{The observation kernel is modeled by $\phi_{4,k}(r-r')$} \dean{this sounds like you are saying that we used a field basis function as the observation kernel. This is hard to justify, because no one will expect the field to be directly related to the sensors. Is it just a coincidence they are the same width? If so, why not just say we used a B-spline function with width ---? Maybe I am missing something.}, providing a width of 0.8 mm at half the maximum amplitude of the observation kernel, allowing for some cross talk in the recordings. The actual connectivity kernel and its decomposition are plotted in \figurename{\ref{fig:KernelAndFreqResponse}} (a).
